\thesistitle

### 2019-06-04 sections/research.tex
Update research from sattose paper

\chapter{Redefining type 1, 2 and type 3 clones}
In the background section we discussed the different types of clones, as defined by Roy et al \cite{roy2007survey}. However, when detecting clones by these definitions, we will detect many clones that are not immediately useful for refactoring. In this chapter, we will discuss the shortcoming of each of the clone types, and propose solutions for them.

\section{Type 1 clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. See figure \ref{fig:type1noclone} for an example of a type 1 clone that is not actually identical. In this figure, you can see how all objects in the supposedly cloned fragment (see the marked fragment) refer to different imports. Because of this, it seems like a clone, but might actually not be. Even the showImage method might refer to a different method (overloaded variants). Many (mainly textual and token-based) clone detectors will flag this as a type 1 clone.

@@

@@

\begin{itemize}
  \item \textbf{Considering types:} The definition of type 2 clones states that types should not be considered. We disagree because types can make a significant change to the meaning of a code segment and thus whether this segment should be considered a clone.
  \item \textbf{Having a distinction between different variables:} By the definition of type 2 clones, any identifiers would not be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \textit{``myMethod(var1, var2)''}, or call this method like this: \textit{``myMethod(var1, var1)''}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} By the definition of type 2 clones any literals would not be taken into account. We agree, as when merging the clone (for instance by extracting a method), we can simply turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
  \item \textbf{Consider method call signatures and define a threshold for variability in method calls:} As type-2 clones allow changes in identifiers, also the names of called methods may vary. However, because of this, completely different methods can be called in cloned fragments as a result. This poses serious challenges on refactoring and makes it more disputable whether such a clone is actually harmful. This is because different method identifiers can describe a completely different functionality. Because of that, we recommend to keep considering the call signatures of cloned methods when they are compared. We can allow variability in the rest of method identifiers by passing the function as a parameter. To limit the amount of parameters required we also recommend defining a threshold for variability in method call expressions, so only a limited number of method calls can vary.

@@

\chapter{Merging duplicate code through refactoring}
Now we have mapped the contexts in which clones occur, we can start looking at refactoring opportunities. Regarding refactoring, we separate clones in two categories: easy and difficult refactoring opportunities. Easy refactoring opportunities are clones that can easily be automatically refactored. Examples of these opportunities are fully cloned methods or a set of fully cloned statements. According to the relation between the clone instances, we can propose a refactoring automatically.

However, not always can a clone easily be merged. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain a complex control structure, like labels, return, break, continue, etc. In such instances, more conditions apply to be able to conduct a refactoring, if advisable at all.

The most trivial way to merge a clone is through method extraction. However, method extraction is not always possible. For instance, if a part of a subtree (in a programs' AST) is matched as a clone. Because of this, we chose to measure what percentage of ``Partial Method'' clones are refactorable using method extraction. Our results are displayed in table \ref{table:refactorability}.

\begin{table}[H]
  \begin{center}
  \caption{Clone instance contents} \label{table:refactorability}
  \bigskip
\begin{tabular}{|l|l|l|}
\hline
\textbf{Refactorability}        & \textbf{Amount} & \textbf{Percentage} \\ \hline
\begin{tabular}[c]{@{}l@{}}Cannot directly\\be extracted\end{tabular}    & 10,990           & 41.91               \\ \hline
\begin{tabular}[c]{@{}l@{}}Is not a\\partial method\end{tabular} & 9,444            & 36.01               \\ \hline
\begin{tabular}[c]{@{}l@{}}Can be\\extracted\end{tabular}                & 5,791            & 22.08               \\ \hline

\end{tabular}
\end{center}
\end{table}

From this table, we can see that approximately 22 percent of the clones can be refactored through method extraction. For the other clones, we must first build a catalog of appropriate refactorings.

>>>>

\chapter{Addressing problems with clone type definitions}
In section \ref{chap:backgroundclonetypes} we discussed the different types of clones, as defined by Roy et al \cite{roy2007survey}. For each of type 1-3 clones we list our solutions to their shortcomings to increase the chance that we can merge the clone while improving the design. Due to the serious challenges involved in their detection and refactoring, type 4 clones are not considered in this study.

\section{Type 1 clones}
\label{chap:type1clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, it does not yet indicate that they are functionally identical. Although textually equal, method calls can still refer to different methods, type declarations can still refer to different types and even variables can be of a different type. In such cases refactoring opportunities could be invalidated. Therefore, we propose an alternative definition of type 1 clones. We have named it ``type 1R clones''. The ``R'' stands for refactoring-oriented (and may be less suitable for other analyses). We describe how type 1R clones differ from type 1 clones over the following paragraphs.

@@

\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are functionally equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality of the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
  \item \textbf{Compare the equality of the fully qualified identifier for method references and call signatures.} This is to prevent a change in functionality after merging the clone, as methods with equal names/parameters can still contain different functionality.
\end{itemize}

@@

\begin{itemize}
  \item \textbf{Considering types:} The definition of type 2 clones states that types should not be considered. We disagree because types can make a significant change to the meaning of a code segment and thus whether this segment should be considered a clone.
  \item \textbf{Having a distinction between different variables:} By the definition of type 2 clones, any identifiers would not be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \textit{``myMethod(var1, var2)''}, or call this method like this: \textit{``myMethod(var1, var1)''}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} By the definition of type 2 clones any literals would not be taken into account. We agree, as when merging the clone (for instance by extracting a method), we can simply turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
  \item \textbf{Consider method call signatures and define a threshold for variability in method calls:} As type-2 clones allow changes in identifiers, also the names of called methods may vary. However, because of this, completely different methods can be called in cloned fragments as a result. This poses serious challenges on refactoring and makes it more disputable whether such a clone is actually harmful. This is because different method identifiers can describe a completely different functionality. Because of that, we recommend to keep considering the call signatures of cloned methods when they are compared. We can allow variability in the rest of method identifiers by passing the function as a parameter. To limit the amount of parameters required we also recommend defining a threshold for variability in method call expressions, so only a limited number of method calls can vary.
\end{itemize}

@@

\chapter{Merging duplicate code through refactoring}
The most used technique to merge clones is method extraction (creating a new method on basis of the contents of clones). However, method extraction cannot be applied in all cases. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain statements like \texttt{return}, \texttt{break}, \texttt{continue}. In these instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all.

We measured the amount of clones that can be refactored through method extraction (without additional transformations being required). Our results are displayed in Table \ref{table:refactorability}. In this table we use the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to merge the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table \ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}

\begin{table}[H]
  \begin{center}
  \caption{Refactorability through method extraction} \label{table:refactorability}
  \medskip
\begin{tabular}{|l|l|l|l|l|}
\hline
\textbf{}        & \textbf{\# T1} & \textbf{\% T1}  & \textbf{\# 1R} & \textbf{\% 1R} \\ \hline
\begin{tabular}[c]{@{}l@{}}Is not a\\partial method\end{tabular}           & 5,917 & 34.22 & 4,806 & 38.49 \\ \hline
\begin{tabular}[c]{@{}l@{}}Complex\\control flow\end{tabular} & 5,511 & 31.87 & 3,158 & 25.29 \\ \hline
\begin{tabular}[c]{@{}l@{}}Spans part of\\a block\end{tabular}             & 3,989 & 23.07 & 3,152 & 25.24 \\ \hline
\begin{tabular}[c]{@{}l@{}}Can be\\extracted\end{tabular}                  & 1,874 & 10.84 & 1,371 & 10.98 \\ \hline
\end{tabular}
\end{center}
\end{table}

From Table \ref{table:refactorability}, we can see that approximately ten percent of the clones can directly be refactored through method extraction (and possibly other refactoring techniques based on the relation of the clone instances). For the other clones, other techniques or transformations will be required. Looking into these techniques and transformations will be one of our next steps.
